When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph Databases

### 2018-02-19 paper/paper.tex
Avoid bad hyphenation
Are real world instances hard (review 3)

In \cref{section:induced} we look at the induced isomorphism problem.  For certain choices of
parameters, there are two phase transitions, going from satisfiable to unsatisfiable, and then from
unsatisfiable back to satisfiable. Again, when going from satisfiable to unsatisfiable (from either
direction), instances go from being trivial to really hard to solve. However, each of the three
solvers we test also finds the central unsatisfiable region to be hard, despite it not being near a
phase transition. To show that this is not a simple weakness of current subgraph isomorphism
algorithms, we verify that this region is also hard for boolean satisfiability, pseudo-boolean, and
mixed integer solvers, and under reduction to the clique problem. Interestingly, the constrainedness
measure proposed by \citet{DBLP:conf/aaai/GentMPW96} \emph{does} predict this difficult region---the
third contribution of this paper is to use these instances to provide evidence in favour of
constrainedness, rather than proximity to a phase transition, being an accurate predictor of
difficulty, and to show that constrainedness is not simply a refinement of a phase transition
prediction.

@@

When labels were introduced, we saw that VF2 finds some instances hard which other algorithms find
easy. We looked at this kind of instance in more detail, and argued that this was due to shortcomings in
VF2's design, rather than an interesting property of \NP-completeness. Although inevitably there
will be instances that every algorithm finds hard, we see it as a serious weakness for an algorithm to exhibit
exponential behaviour in such trivial cases as when there are two red vertices in the pattern but only one in the
target.

>>>>

In \cref{section:induced} we look at the induced subgraph isomorphism problem.  For certain choices of
parameters, there are two phase transitions, going from satisfiable to unsatisfiable, and then from
unsatisfiable back to satisfiable. Again, when going from satisfiable to unsatisfiable (from either
direction), instances go from being trivial to really hard to solve. However, each of the three
solvers we test also finds the central unsatisfiable region to be hard, despite it not being near a
phase transition. To show that this is not a simple weakness of current subgraph isomorphism
algorithms, we verify that this region is also hard for boolean satisfiability, pseudo-boolean, and
mixed integer solvers, and under reduction to the clique problem. Interestingly, the constrainedness
measure proposed by \citet{DBLP:conf/aaai/GentMPW96} \emph{does} predict this difficult region---the
third contribution of this paper is to use these instances to provide evidence in favour of
constrainedness, rather than proximity to a phase transition, being an accurate predictor of
difficulty, and to show that constrainedness is not simply a refinement of a phase transition
prediction.

@@

We note that these ``really hard'' instances have a very particular structure, which is unlikely to
arise naturally. Indeed, the successes seen with constraint programming approaches on the instances
in \cref{section:filterverify}, and in the wider range of application instances used by
\citet{DBLP:conf/lion/KotthoffMS16}, suggest that the real-world instances encountered so far are
generally fairly easy for a good algorithm. Nonetheless, carefully designed synthetic instances
provide a good starting point for scientific experiments, and can help improve the behaviour of
algorithms on real-world instances too.

For example, when labels were introduced, we saw that VF2 found some instances hard which other
algorithms found easy. We looked at this kind of instance in more detail, and argued that this was
due to shortcomings in VF2's design, rather than an interesting property of \NP-completeness.
Although inevitably there will be instances that every algorithm finds hard, we see it as a serious
weakness for an algorithm to exhibit exponential behaviour in such trivial cases as when there are
two red vertices in the pattern but only one in the target.
